Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2019-05-25 papers/Visual Diagnostics of a Model Explainer/visual-diagnostics-of-a-model-explainer.Rnw
updated intro
worked more on intro to paper - getting there...

\documentclass[AMS,STIX1COL]{WileyNJD-v2}

@@

Machine learning algorithms are known to provide accurate predictions. (include more about the usefulness of ml algorithms). However, this predictive ability comes at a cost. These algorithms are typically very complex and have no functional form. As a result, the ability to directly interpret the model is lost.

In many areas of application, the need to be able to explain the results from the model is critical. For example, in forensics science a machine learning model could be used to determine the probability that a shoe print found at the crime scene matches the shoe of a suspect. The model may return a high probability of a match, but without the ability to explain which factors in the model led to this high probability, it may be difficult for a jury to trust the results from the model when deciding whether or not to convice the suspect. (add sentence about Europe law) This has led to the desire to be able to interpret machine learning algorithms.

One approach to remedy this problem would be to develop a model explainer that can provide insight into the model performance. A model explainer is a method that is separate from the model but uses the model to provide understanding into mechanism of how the model produces predictions. (try to think of a simple example of a model explainer)

Locally interpretable model explanations (LIME) is one such model explainer (reference). While some model explainers are focused on understanding a model at the global level, LIME was designed to explain the model for a single prediction of interest. Conceptually, LIME fits a simple model that is able to be interpreted, which is referred to as the explainer, that is meant to capture the behavior of the complex "black box model" in a local region around a prediction of interest. (a good place to include an image to help with this explanation)

@@

@@

- explanation is also a prediction
  - how reliable is that explanation?
- we will present some tools that could help answer that question
- we will show an example using this tools (bullet data)
- we will mention some possible insights about LIME that these tools have shown us
1. explainer model generally has very low $R^2$ (probably due to binning)
2. "local" explanations are not local but are driven by the ("global") marginal distributions of covariates 
- Describe LIME including details on binning and linear regression in binned features and motivation for the binning

@@

>>>>

\documentclass[AMS,STIX2COL]{WileyNJD-v2}

@@

Machine learning models have been proven to provide accurate predictions in many areas of applications. Carriquiry et. al. (source) discuss improvement in various pattern recognition analyses in the forensics sciences with the use of random forests and neural networks. (need a source) describe the numerous ways machine learning has been used in the medical field to advance healthcare. However, this predictive ability comes at a cost.

These models are typically complex and have no simple functional form. As a result, the ability to directly interpret the model is lost. In many areas of application such as with forensics science and healthcare, the need to be able to explain the results from the model is critical. For example, a machine learning model could be used to determine the probability that a shoe print found at the crime scene matches the shoe belonging to a suspect in the investigation. The model may return a high probability of a match, but without the ability to explain which factors in the model led to this high probability, it may be difficult for a jury to trust the results from the model when deciding whether to convict the suspect. (add sentence about Europe law) The importance of interpretability has led to the desire to be able to explain the output from machine learning algorithms.

One approach to remedy this problem would be to develop a model explainer that can provide insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model to provide understanding into mechanism of how the model produces predictions. (try to think of a simple example of a model explainer) (mention some of the model explainers that have been developed recently)

LIME (local interpretable model explanations) is one such model explainer (reference). While some model explainers are focused on understanding a model at the global level, LIME was designed to explain the model at the local level for a single prediction of interest. Conceptually, LIME fits a simple model that can be interpreted, referred to as the explainer, that is meant to capture the behavior of the complex "black box model" in a local region around the prediction of interest. (a good place to include an image to help with this explanation)

- explanation is also a prediction/model
  - how reliable is that explanation?
  - also, no current recommendations for LIME algorithm input options

@@

- we will present some tools that could help answer that question
- we will show an example using this tools (bullet data)
- we will mention some possible insights about LIME that these tools have shown us

\subsection{Overview of LIME}

@@

@@

1. explainer model generally has very low $R^2$ (probably due to binning)
2. "local" explanations are not local but are driven by the ("global") marginal distributions of covariates
